R version used to run this code: R 4.4.1


```{r}
# # Below code can be used to install required dependencies
# # Install dependency
# if (!require("BiocManager", quietly = TRUE))
#   install.packages("BiocManager")
# 
# BiocManager::install("STexampleData")
# install.packages("devtools")


# # Installing waveST
# library(devtools)
# install_github("OliverXUZY/waveST")
```

```{r}
library(waveST)
library(fields)
library(tidyverse)
theme_set(theme_minimal())
```

```{r}
set.seed(100) # For reproducibility
```

## Data simulation

Retrieved and modified from: https://github.com/OliverXUZY/waveST/blob/master/vignettes/vignette01.Rmd

```{r}
########## ==================  Data Generation
NUM_GENES = 300
K = 9

## == Step 1. generate coordinate matrix image with different patterns
# @N_dup, generate duplicates of each matrix pattern with a little permutation

CMls = generateCoorM(1)
```


We show all the generated pattern (the colours may differ from the one from the waveST paper):
```{r, fig.width= 8, fig.height=8, result = "hide"}
# show in scaled version
lay = layout(matrix(1:9, 3, 3, byrow = TRUE))
CMls %>%
  map(~ .x / sqrt(sum(.x^2))) %>%
  map(~ image.plot(.x, asp = 1))
layout(1)
```

We treat previous 9 patterns as factor genes, i.e. we have 8 factors matrix F = D^2xK (1024x8), where 1024 comes from (D^2) 32^2. We consider p = 500 genes in this simulation generate nxp(1024x500) data matrix by F%*%V, where V is \code{K}xp (8x500), compute the magnitude of each image matrix

```{r}
scales = CMls %>%
  map_dbl(~ sqrt(sum(.x^2)))
## rescale each image matrix to norm 1 (eigen gene has norm 1)
FctGenes = CMls %>%
  map(~ .x / sqrt(sum(.x^2))) %>%
  map_dfc(~ as.vector(.)) %>%
  as.matrix()

LoadGenes = rep(1000, 9) %>%
  map_dfc(~ .x * rnorm(NUM_GENES)) %>%
  as.matrix(.) %>%
  t(.)

truth = FctGenes %*% LoadGenes

## we generate data by add random noise
raws = truth + matrix(rnorm(prod(dim(truth)), mean = 0, sd = 190), nrow = dim(truth)[1], ncol = dim(truth)[2])

## we standardised data to 0 mean and 1 standard deviation (for consistency with the WaviFM prepreocessing)
std_raws = scale(raws, center=TRUE, scale=TRUE)

## construct waveST object
wave = waveST(data = std_raws)
```

## Exporting data

```{r}
write.csv(std_raws, 'std_raws.csv')
```

```{r}
vector_list <- list()

for (k in 1:K) {
  vector_list[[k]] <- CMls[[k]] %>% as.matrix() %>% as.vector()
}

true_factors <- do.call(cbind, vector_list)
write.csv(true_factors, "true_factors.csv")
```

## Perform waveST

Retrieved and modified from: https://github.com/OliverXUZY/waveST/blob/master/vignettes/vignette01.Rmd


```{r}
# Uses wavelet transform, manual thresholding with various threshold levels, SVD factor model
decomp_0_threshold = decompose(wave, "wave", "SVD", K = K, J=3, thresholdMethod="manual", tau=0)
decomp_0_threshold_5 = decompose(wave, "wave", "SVD", K = K, J=5, thresholdMethod="manual", tau=0)
```


```{r}
# Uses wavelet transformation, hybrid thresholding, SVD factor model
decomp_hybrid = decompose(wave, "wave", "SVD", K = K, J=3, thresholdMethod="hybrid")
decomp_hybrid_5 = decompose(wave, "wave", "SVD", K = K, J=5, thresholdMethod="hybrid")
```

## Export inferred factor activites

```{r}
save_waveST_L <- function(decomp, output_csv, K) {
  vector_list <- list()
  
  for (k in 1:K) {
    vector_list[[k]] <- InvWaveTrans(decomp@output$f[, k], decomp@input %>% as.matrix()) %>% as.vector()
  }
  
  f_matrix <- do.call(cbind, vector_list)
  
  # Export the matrix to a CSV file
  write.csv(f_matrix, output_csv)
}
```


```{r}
save_waveST_L(decomp_0_threshold, "waveST0Threshold.csv", K)
save_waveST_L(decomp_0_threshold_5, "waveST0Threshold5.csv", K)
save_waveST_L(decomp_hybrid, "waveSTHybrid.csv", K)
save_waveST_L(decomp_hybrid_5, "waveSTHybrid5.csv", K)
```


